Dependent Object Types

### 2012-08-06 article.tex
remote template blurb
fix fun sugar
make todos and remarks explicit
grammar
sketch instead of prove

%-----------------------------------------------------------------------------
%
%               Template for sigplanconf LaTeX Class
%
% Name:         sigplanconf-template.tex
%
% Purpose:      A template for sigplanconf.cls, which is a LaTeX 2e class
%               file for SIGPLAN conference proceedings.
%
% Guide:        Refer to "Author's Guide to the ACM SIGPLAN Class,"
%               sigplanconf-guide.pdf
%
% Author:       Paul C. Anagnostopoulos
%               Windfall Software
%               978 371-2316
%               paul@windfall.com
%
% Created:      15 February 2005
%
%-----------------------------------------------------------------------------

@@

\title{Dependent Object Types}
\subtitle{A foundations for Scala's type system} % TODO: ask

@@

We show that DOT and its patched-up variants are not syntactically
sound, by exhibiting counterexamples to preservation. Nevertheless, we
prove type-safety of the calculus via step-indexed logical relations.
\end{abstract}

@@

% TODO: ask: does this last point still makes sense wrt to type safety via logical relations instead of via the standard theorems of preservation and progress?

@@

\subsection{Sugar: Functions}
Like in Scala, we can encode functions as objects with a special
method. We will freely use the following sugar in the remaining of
this paper. Note that the variable $z$ must be fresh.
\begin{align*}
\arrow S T &\iff {\Top \refine z {\mdecl {\mi{apply}} S T}}\\
\fun x S T t &\iff \new z {\arrow S T} \ldefs{\mi{apply}(x) = t}\\
\app f x &\iff \mapp f {\mi{apply}} x\\
\cast T t&\iff \app {\fun x T T x} t
\end{align*}

@@

We prove that the DOT calculus is type-safe using logical relations in
section~\ref{dot-type-safety}. However, we first tried to prove the
calculus type-safe using the standard theorems of preservation and
progress~\cite{soundness,tapl}. Unfortunately, for the calculus as
presented, and any of its variants that we devised, preservation
doesn't hold. In this section, we review some of the most salient
counterexamples to preservation that we found. These counterexamples
have been checked with PLT Redex~\cite{plt_redex}.

@@

$a.i.l$ reduces to $b.l$. $b.l$ has type $b.X$, so we need $b.X <:
a.i.X$. This cannot be established with the current rules: it is not
true in general, but true here because $a.i$ reduces to $b$. Hence,
the need for acknowledging path equality for preservation to hold.

\section{Type Safety via Logical Relations}\label{dot-type-safety}
We prove type-safety of the DOT calculus via step-indexed logical
relations~\cite{ahmed04,ahmed06,step_indexed_obj}.

% TODO: section

@@

% TODO: more discussion subsections?

\section{Conclusion}\label{conclusion}
% TODO: section

>>>>

@@

\title{Dependent Object Types}
\subtitle{Towards a foundation for Scala's type system}

@@

We show that DOT and its patched-up variants are not syntactically
sound, by exhibiting counterexamples to preservation. Nevertheless, we
sketch a proof of type-safety of the calculus via step-indexed logical
relations.
\end{abstract}

@@

\remark{Does this last point still makes sense wrt to type safety via
  logical relations instead of via the standard theorems of
  preservation and progress?}

@@

\subsection{Sugar: Functions}
Like in Scala, we can encode functions as objects with a special
method. We will freely use the following sugar in the remaining of
this paper. Note that the variable $z$ must be fresh.
\begin{align*}
\arrow S T &\iff {\Top \refine z {\mdecl {\mi{apply}} S T}}\\
\fun x S T t &\iff \new z {{\arrow S T}\ldefs{\mi{apply}(x) = t}} z\\
\app f x &\iff \mapp f {\mi{apply}} x\\
\cast T t&\iff \app {\fun x T T x} t
\end{align*}

@@

We sketch a proof that the DOT calculus is type-safe using logical
relations in section~\ref{dot-type-safety}. However, we first tried to
prove the calculus type-safe using the standard theorems of
preservation and progress~\cite{soundness,tapl}. Unfortunately, for
the calculus as presented, and any of its variants that we devised,
preservation doesn't hold. In this section, we review some of the most
salient counterexamples to preservation that we found. These
counterexamples have been checked with PLT Redex~\cite{plt_redex}.

@@

$a.i.l$ reduces to $b.l$. $b.l$ has type $b.X$, so we need $b.X <:
a.i.X$. This cannot be established with the current rules: it is not
true in general, but true here because $a.i$ reduces to $b$. Hence,
we need to acknowledge path equality for preservation to hold.

\section{Type Safety via Logical Relations}\label{dot-type-safety}
We sketch a proof of type-safety of the DOT calculus via step-indexed
logical relations~\cite{ahmed04,ahmed06,step_indexed_obj}.

\todo{complete proof!}

@@

\todo{more discussion subsections?}

\section{Conclusion}\label{conclusion}
\todo{conclusion}
